The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-06-12 guides/practical_guides.Rmd
guides; drake -> targets

---
title: "Practical guides for implementing 'Five Pillar' principles"
author: "Mark Ziemann, Samuel S. Minot, Pierre Poulain, Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: practical_guides.bib
csl: plos-computational-biology.csl
---

Mark Ziemann<sup>1*</sup>, Samuel S. Minot<sup>2</sup>, Pierre Poulain<sup>3</sup>,
Anusuiya Bora<sup>1</sup>

<details><summary>Author information (*click to expand*)</summary><p>

@@

2. Fred Hutchinson Cancer Center, Seattle, USA.

3. Université Paris Cité, CNRS, Institut Jacques Monod, Paris, France.

@@

| Author | ORCID |
| --- | --- |
| Mark Ziemann | 0000-0002-7688-6974 |
| Samuel S. Minot | 0000-0003-1639-3905 |
| Pierre Poulain | 0000-0003-4177-3619 |
| Anusuiya Bora | 0009-0006-2908-1352 |

@@

Source: https://github.com/markziemann/5pillars/blob/main/manuscript/practical_guides.Rmd

@@

@@

* [Ebook:Happy Git and GitHub for the useR](https://happygitwithr.com/index.html)

@@

* [Guide: Introduction to data sharing](https://zenodo.org/badge/DOI/10.5281/zenodo.3518692.svg)

@@

@@

* [Tutorial: A short introduction to drake (An R based workflow manager)](https://data.nozav.org/posts/2019-a-short-introduction-to-drake/)

@@

>>>>

---
title: "Practical guides for implementing 'Five Pillar' principles"
author: "Mark Ziemann, Pierre Poulain, Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: practical_guides.bib
csl: plos-computational-biology.csl
---

Mark Ziemann<sup>1*</sup>, Pierre Poulain<sup>2</sup>,
Anusuiya Bora<sup>1</sup>

Author information

@@

2. Université Paris Cité, CNRS, Institut Jacques Monod, Paris, France.

@@

| Author | ORCID |
| --- | --- |
| Mark Ziemann | 0000-0002-7688-6974 |
| Pierre Poulain | 0000-0003-4177-3619 |
| Anusuiya Bora | 0009-0006-2908-1352 |

@@

Source: https://github.com/markziemann/5pillars/blob/main/guides/practical_guides.Rmd

@@

* [Video: Running Basic Statistical Analysis in R](https://www.youtube.com/watch?v=I6FJo8x1wZE)

@@

* [Ebook: Happy Git and GitHub for the useR](https://happygitwithr.com/index.html)

@@

* [Guide: Introduction to data sharing](https://doi.org/10.5281/zenodo.3518692)

@@

* [Documentation: FAIRsharing.org](https://fairsharing.org/)

<br>

@@

* [Documentation: The {targets} R package user manual (workflow manager for R)](https://books.ropensci.org/targets/)

@@

* [Ebook: Bioinformatics Data Skills Chapter 2](https://vincebuffalo.com/book/)

### 2023-06-12 manuscript/5pillars.Rmd
guides; drake -> targets

A guiding principle of the five pillars approach is that the publishing researchers should strive
to make the reproduction process easier and faster for those who wish to.
A "master script" that coordinates the execution of individual parts of an analysis is an excellent
way to reduce the number of commands required for reproduction.
This can be coded in standard open source languages like R/R Markdown, Python/Jupyter notebook or the
Unix Shell.
Using free and open source software ensures that the research is accessible to the greatest audience
@Garijo2013-zl, as opposed to proprietary software like SPSS, STATA and MatLab which are cost inhibitory.
Free and open source software also enjoy relatively larger libraries of add-on packages contributed by
the scientific community.
Projects entailing big data analysis might benefit from a build/workflow automation solution like `make`,
`snakemake` [@Koster2012-sq; @Molder2021-mp], `drake` [@Michael_Landau2018-kk] or
`nextflow` [@Di_Tommaso2017-yo].
The advantage of such tools is that if the analysis were to terminate due to an issue mid-way,
for example, a hardware problem at step 8 of a 15-step workflow, the analysis of steps 1-7 wouldn't
need to be repeated.
After fixing the hardware issue, re-running the analysis would pick up at step 8 again,
which saves labor and compute time.

>>>>

A guiding principle of the five pillars approach is that the publishing researchers should strive
to make the reproduction process easier and faster for those who wish to.
A "master script" that coordinates the execution of individual parts of an analysis is an excellent
way to reduce the number of commands required for reproduction.
This can be coded in standard open source languages like R/R Markdown, Python/Jupyter notebook or the
Unix Shell.
Using free and open source software ensures that the research is accessible to the greatest audience
@Garijo2013-zl, as opposed to proprietary software like SPSS, STATA and MatLab which are cost inhibitory.
Free and open source software also enjoy relatively larger libraries of add-on packages contributed by
the scientific community.
Projects entailing big data analysis might benefit from a build/workflow automation solution like `make`,
`snakemake` [@Koster2012-sq; @Molder2021-mp], `targets` [@Landau2012] or
`nextflow` [@Di_Tommaso2017-yo].
The advantage of such tools is that if the analysis were to terminate due to an issue mid-way,
for example, a hardware problem at step 8 of a 15-step workflow, the analysis of steps 1-7 wouldn't
need to be repeated.
After fixing the hardware issue, re-running the analysis would pick up at step 8 again,
which saves labor and compute time.
